---
title: "Benchmark analysis for Spatial GRF"
author: "Manh Hung LE"
date: 2026-04-06
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmark analysis for Spatial GRF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup_vignette, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

This vignette presents a benchmark of the LearnerRegrGRF (Geographical Weighted Random Forest) against standard regression models (such as Featureless, CV_Glmnet and KNN).
We evaluate the models using two datasets with spatial coordinates to evaluate our learner.

# Setup and datasets

We are using California Housing (subsampled to 200 points) and Income for our two datasets with spatial coordinates.
The mlr3spatiotempcv package is also used to define these as Spatio-Temporal regression tasks.  

```{r setup_data, message = FALSE, warning = FALSE}
library(mlr3)
library(mlr3spatiotempcv)
library(mlr3learners)
library(mlr3tuning)
library(SpatialML)
library(mlr3learners.spatialML)
library(data.table)
library(ggplot2)
library(future)

task_list <- list()
sample_size_california <- 200
seed_value <- 1

# California Housing Task
california_dt <- mlr3::tsk("california_housing")$data()
california_dt <- california_dt[complete.cases(california_dt), ]

set.seed(seed_value)

california_dt <- california_dt[sample(.N, sample_size_california), ]
california_dt$ocean_proximity <- NULL

task_list[["California"]] <- as_task_regr_st(
  california_dt,
  target = "median_house_value",
  coordinate_names = c("longitude", "latitude"),
  id = "California_Housing"
)

# Income Task
data(Income, package = "SpatialML")
income_df <- as.data.frame(Income)

task_list[["Income"]] <- as_task_regr_st(
  income_df,
  target = "Income01",
  coordinate_names = c("X", "Y"),
  id = "Income"
)
```

# Learners Configuration

As mentionned before, we are comparing four learners: our GRF, a Featureless model, a Linear Model (CV_Glmnet) and a tuned KNN.

```{r setup_learners}
grf_bw <- 20
grf_ntree <- 50

# Learners definition
grf_lrn <- lrn("regr.grf", bw = grf_bw, ntree = grf_ntree, id = "GRF")
featureless_lrn <- lrn("regr.featureless", id = "featureless")
cv_glmnet_lrn <- lrn("regr.cv_glmnet", id = "cv_glmnet")

knn_lrn <- lrn("regr.kknn", id = "knn")
knn_lrn$param_set$values$k <- paradox::to_tune(1, 30)
kfoldcv <- mlr3::rsmp("cv")
kfoldcv$param_set$values$folds <- 3

knn_tuned_lrn <- mlr3tuning::auto_tuner(
  learner = knn_lrn,
  tuner = mlr3tuning::tnr("grid_search"),
  resampling = kfoldcv,
  measure = mlr3::msr("regr.rmse")
)

learners <- list(
  grf_lrn, 
  featureless_lrn, 
  knn_tuned_lrn, 
  cv_glmnet_lrn
)

```

# Benchmark

To evaluate the performance, we run the benchmark locally using the future package for parallel processing.
This allows us to execute the 5-fold cross-validation across all learners and tasks efficiently.

```{r setup_benchmark, message = TRUE, warning = FALSE}
library(future)

cv <- mlr3::rsmp("cv", folds = 5)
bench_grid <- mlr3::benchmark_grid(
  tasks = task_list, 
  learners = learners, 
  resampling = cv
)

if(require(future)) {
  plan("multisession", workers = 4)
}
bench_results <- suppressWarnings(
  mlr3::benchmark(bench_grid, store_models = FALSE)
)

scores_dt <- bench_results$score(mlr3::msr("regr.rmse"))
rm(bench_results)
```

# Results Visualization 
We visualize the Root Mean Squared Error (RMSE) to compare the performance of our learners.
A lower RMSE indicates a better fit to the spatial data.

```{r setup_results}

results_dt <- as.data.table(scores_dt)

results_dt[, algorithm := gsub("regr.", "", learner_id, fixed = TRUE)]
results_dt[, dataset := task_id]
results_dt[, rmse := regr.rmse]

algo_order <- results_dt[, .(m = mean(rmse)), by = algorithm][order(-m), algorithm]
results_dt[, algorithm := factor(algorithm, levels = algo_order)]

results_dt[, dataset := factor(dataset, levels = c("Income", "California_Housing"))]

ggplot(results_dt, aes(x = rmse, y = algorithm)) +
  geom_point(
    shape = 21,
    size = 2,
    fill = "white",
    color = "black",
    stroke = 1
  ) +
  labs(
    x = "RMSE",
    y = "Algorithm"
  )

ggplot(results_dt,  aes(x = rmse, y = algorithm)) +
    geom_point(
    shape = 21,
    size = 2,
    fill = "white",
    color = "black",
    stroke = 1
  ) +
  facet_grid(dataset ~ .) +
  labs(
    x = "RMSE",
    y = "Algorithm"
  )
```

